Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-09 WSSSPE3/report/wg_appendix_funding_spec_expert.tex
Fix typos

The UK allows this kind of collaboration to support the creation of scientific
software for the large supercomputing resource called ARCHER.
However, as well as resourcing this from the staff of the Edinburgh Parallel
Computing Centre, who host the computer, this ``embedded CSE'' resource also
allows the programming to come from local groups. This has been very helpful
in providing funding to establish local groups. These work best when they
develop good collaborations with national cyberinfrastucture pools.
When an organization assembles
a developer pool, diversity is developed and skills can be transferred.

We would like to see these models applied outside High Performance Computing;
most scientific software is not destined to run on national cyberinfrastructure,
but needs similar support. The argument regarding making better use of expensive
hardware through software improvements has been useful politically, (and many
RSE groups are cited in organisations which host clusters for this reason), but
the time has come to make the case that software itself is a critical
cyberinfrastructure, and, with a much longer shelf-life than hardware, is itself
a capital investment.

@@

We think support from non-governmental organisations will be important if RSE
groups will become established.
The Sloan Foundation is currently funding data science engineers, who work in
the context of other SW developers at University of Washington.
These scientists work in the e-Science Studio/dataScience Studio, and they
help a group of graduate students in solving their problems in data
science and data management. During Fall and Spring, a 10-week incubator program
allows students to work two days a week to work on a data-intensive science
project. Some fraction of the developer time is dedicated to the developers'
personal interests as well as instruction.

@@

One model that has been helpful in finding funding for RSE groups is the use of
funds left over on research grants when RAs have left prematurely -- PIs like
this arrangement as it is hard to find good staff for short-term positions, so
having a pool of research programming staff on hand resolves this problem.
We recommend that funders give explicit guidance to grant holders and
institutions that such arrangement are favourable. Framework agreements
permitting this to go ahead without
checking back every time with funders and/or grant panels would further
smooth this.
(This is also provides a more stable job for those who hold these skills, but
arguments about making life nicer for postdocs will not help persuade funders
or PIs!)

@@

There is as yet no clear answer as to the scale of aggregation needed to
make such a program work. A university wide programme allows enough scale to
be robust to fluctuations of funding within one field.
But a specialization focus on developers to support, for example,
physical or biological
sciences may be preferable, if the customer base is large enough.
The desire to aggregate enough work to make it sustainable, and the need to
have domain-relevant research programming skils, are in tension.

In the UK, another source of funding for research software has been the
Collaborative computational projects (CCPs): domain specific communities put
forward proposals that are a priority of the community as a whole, for example,
biosimulation or plasma physics. These bodies act as custodians of community codes,
and a central team also provides software engineering support support.

However this area develops, the need for funding to secure software as a
cyberinfrastructure component is clear. Funding which permits code to be
refactored, tidied and optimised is rare; this is often done ``on the sly''
in a scientifically focused grant. The UK EPSRC's ``software for the future''
call, which really permits explicit investiment in software as an infrastructure,
is so oversubscribed as to have a FOUR PERCENT success rate; the demand is clear!

@@

We will seek to identify and approach existing research programming organisations,
to get their permission to list them on a list of research software groups.
We will also look for examples of groups which have successfully become self-
sustaining following

@@

The UKRSE community will provide initial facilities to host this list, and
continue to work to spread the initiative, but local leadership in the US
is needed if this campaign is to succeed. This will require an initial gathering
of identified research software organisations in the US to this end.

@@

Financial support for an initial conference to bring together research software
groups to form an organisation and create a resource sharing structure would
help to further this campaign.

In the longer term, funding organisations, especially non-governmental organisations
with the capability to effect innovation through seed funding, could provide
support to nucleate the creation of research software groups.

>>>>

The UK allows this kind of collaboration to support the creation of scientific
software for the large supercomputing resource called ARCHER.
However, as well as resourcing this from the staff of the Edinburgh Parallel
Computing Centre, who host the computer, this ``embedded CSE'' resource also
allows the programming to come from local groups. This has been very helpful
in providing funding to establish local groups. These work best when they
develop good collaborations with national cyberinfrastructure pools.
When an organization assembles
a developer pool, diversity is developed and skills can be transferred.

We would like to see these models applied outside High Performance Computing;
most scientific software is not destined to run on national cyberinfrastructure,
but needs similar support. The argument regarding making better use of expensive
hardware through software improvements has been useful politically, (and many
RSE groups are cited in organizations which host clusters for this reason), but
the time has come to make the case that software itself is a critical
cyberinfrastructure, and, with a much longer shelf-life than hardware, is itself
a capital investment.

@@

We think support from non-governmental organizations will be important if RSE
groups will become established.
The Sloan Foundation is currently funding data science engineers, who work in
the context of other SW developers at University of Washington.
These scientists work in the e-Science Studio/dataScience Studio, and they
help a group of graduate students in solving their problems in data
science and data management. During Fall and Spring, a 10-week incubator program
allows students to work two days a week to work on a data-intensive science
project. Some fraction of the developer time is dedicated to the developers'
personal interests as well as instruction.

@@

One model that has been helpful in finding funding for RSE groups is the use of
funds left over on research grants when RAs have left prematurely -- PIs like
this arrangement as it is hard to find good staff for short-term positions, so
having a pool of research programming staff on hand resolves this problem.
We recommend that funders give explicit guidance to grant holders and
institutions that such arrangement are favorable. Framework agreements
permitting this to go ahead without
checking back every time with funders and/or grant panels would further
smooth this.
(This is also provides a more stable job for those who hold these skills, but
arguments about making life nicer for postdocs will not help persuade funders
or PIs!)

@@

There is as yet no clear answer as to the scale of aggregation needed to
make such a program work. A university wide program allows enough scale to
be robust to fluctuations of funding within one field.
But a specialization focus on developers to support, for example,
physical or biological
sciences may be preferable, if the customer base is large enough.
The desire to aggregate enough work to make it sustainable, and the need to
have domain-relevant research programming skills, are in tension.

In the UK, another source of funding for research software has been the
Collaborative computational projects (CCPs): domain specific communities put
forward proposals that are a priority of the community as a whole, for example,
biosimulation or plasma physics. These bodies act as custodians of community codes,
and a central team also provides software engineering support.

However this area develops, the need for funding to secure software as a
cyberinfrastructure component is clear. Funding which permits code to be
refactored, tidied and optimized  is rare; this is often done ``on the sly''
in a scientifically focused grant. The UK EPSRC's ``software for the future''
call, which really permits explicit investment in software as an infrastructure,
is so oversubscribed as to have a FOUR PERCENT success rate; the demand is clear!

@@

We will seek to identify and approach existing research programming organizations,
to get their permission to list them on a list of research software groups.
We will also look for examples of groups which have successfully become self-
sustaining following

@@

The UKRSE community will provide initial facilities to host this list, and
continue to work to spread the initiative, but local leadership in the US
is needed if this campaign is to succeed. This will require an initial gathering
of identified research software organizations in the US to this end.

@@

Financial support for an initial conference to bring together research software
groups to form an organization and create a resource sharing structure would
help to further this campaign.

In the longer term, funding organizations, especially non-governmental organizations
with the capability to effect innovation through seed funding, could provide
support to nucleate the creation of research software groups.
